APPM 3310 Project\\\Large{Static Analysis of Truss Systems}

### 2016-04-25 paper.tex
Write globalization section.

By making the simplification that nodes cannot impart any torque on elements
the only equation governing an element is Hook's law.  In the element centered
coordinate frame the one dimensional Hook's law, Eq.~(\ref{eq:hooks_law}) is
enough.  However, in order to convert this into the global coordinate frame a
full dimensional equation is needed.  This equation will be designated as the
local stiffness equation:
\begin{equation}
    \overline{\mathbf{f}}=\overline{K}\overline{\mathbf{u}}
    \label{eq:local_stiffness_equation}
\end{equation}
This equation is in a member centric coordinate frame with the origin placed
on the first node of the element and the $x$-axis along the element, pointing
to the 2nd node.  The direction of all other axes are unimportant.  Therefore,
in the 2-dimensional case $\overline{K}$ becomes:
\begin{equation}
    \overline{K}=\frac{EA}{L}\begin{bmatrix*}[r]
        1&0&-1&0\\
        0&0&0&0\\
        -1&0&1&0\\
        0&0&0&0
    \end{bmatrix*}
\end{equation}
The 3-dimensional case can be easily obtained by adding 2 zero rows and 2 zero
columns.

@@

Before the global stiffness matrix can be formed
Eq.~(\ref{eq:local_stiffness_equation}) must be converted into the global
coordinate frame equation:
\begin{equation}
    \mathbf{f}^e=K^e\mathbf{u}^e
\end{equation}
which relates displacement to force for a single element $e$ in the global
coordinate frame.  Eq.~(\ref{eq:global_to_local}) shows the relation between the
global coordinate frame and the local coordinate frame.
\begin{equation}
    \overline{\mathbf{u}}^e=T^e\mathbf{u},\qquad
    \overline{\mathbf{f}}^e=T^e\mathbf{f}
    \label{eq:global_to_local}
\end{equation}

>>>>

By making the simplification that nodes cannot impart any torque on elements
the only equation governing an element is Hook's law.  In the element centered
coordinate frame the one dimensional Hook's law, Eq.~(\ref{eq:hooks_law}) is
enough.  However, in order to convert this into the global coordinate frame a
full dimensional equation is needed.  This equation will be designated as the
local stiffness equation for element $e$:
\begin{equation}
    \overline{\mathbf{f}}^e=\overline{K}^e\overline{\mathbf{u}}^e
    \label{eq:local_stiffness_equation}
\end{equation}
This equation is in a member centric coordinate frame with the origin placed
on the first node of the element and the $x$-axis along the element, pointing
to the 2nd node.  The direction of all other axes are unimportant.  Therefore,
in the 2-dimensional case $\overline{K}^e$ becomes:
\begin{equation}
    \overline{K}^e=\frac{EA}{L}\begin{bmatrix*}[r]
        1&0&-1&0\\
        0&0&0&0\\
        -1&0&1&0\\
        0&0&0&0
    \end{bmatrix*} \label{eq:kbar}
\end{equation}
The 3-dimensional case can be easily obtained by adding 2 zero rows and 2 zero
columns.

@@

Before the global stiffness matrix can be formed
Eq.~(\ref{eq:local_stiffness_equation}) must be converted into the global
coordinate frame equation:
\begin{equation}
    \mathbf{f}^e=K^e\mathbf{u}^e\label{eq:global_stiffness_equation}
\end{equation}
which relates displacement to force for a single element $e$ in the global
coordinate frame.  Let $\mathbf{x}$ be a vector in the global coordinate
system and $\overline{\mathbf{x}}$ be the equivalent vector in the element
local coordinate frame.  Then the relation between the two is given by
\parencite{olver2006}:
\begin{equation}
    S\overline{\mathbf{x}}=\mathbf{x}\qquad\text{where}\qquad
    S=\left\{\mathbf{v}_1,\mathbf{v}_2,\ldots,\mathbf{v}_n\right\}
\end{equation}
With the $\mathbf{v}$ vectors forming a basis for the element centered
coordinate system.  In particular, $v_1$ is the unit vector along the element.
If the other columns of $S$ are chosen such that $S$ is orthonormal then
$S^{-1}=S^{T}$.  Therefore, the reverse transformation is:
\begin{equation}
    \overline{\mathbf{x}}=S^{-1}\mathbf{x}=S^T\mathbf{x}
\end{equation}
Noting that $\mathbf{u}$ and $\mathbf{f}$ are stacked vectors that rotate
though each coordinate twice the appropriate transformation matrix for element
$e$ becomes:
\begin{equation}
    T^e=\begin{pmatrix}
        S^T&O\\O&S^T
    \end{pmatrix}
\end{equation}
Therefore, Eq.~(\ref{eq:global_to_local}) shows the relation between the
global coordinate frame and the local coordinate frame for the force and
displacement vectors.
\begin{equation}
    \overline{\mathbf{u}}^e=T^e\mathbf{u},\qquad
    \overline{\mathbf{f}}^e=T^e\mathbf{f}
    \label{eq:global_to_local}
\end{equation}
By combining Eq.~(\ref{eq:local_stiffness_equation},
\ref{eq:global_stiffness_equation}, \ref{eq:global_to_local}) the global
stiffness matrix (for element $e$), $K^e$ can be related to the local
stiffness matrix $\overline{K}^e$ by:
\begin{equation}
    K^e=(T^e)^T\overline{K}^eT^e
\end{equation}
